Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-06 paper.qmd
Added another projection

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

@@

trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
  plot_layout(guides = "collect", nrow = 2) &
  theme(legend.position='none')
```

>>>>

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)

## Second projection
projection <- cbind(
  c(0.30943,-0.08452,-0.23005,-0.05284),
  c(0.22463,-0.06682,0.32106,0.02452))

proj_obj2 <- get_projection(projection = projection, 
                            proj_scale = 3, 
                            scaled_data = scaled_gau_data, 
                            scaled_data_model = scaled_gau_data_model, 
                            distance_df_small_edges = distance_df_small_edges, 
                            axis_param = list(limits = 1.2,
                                              axis_scaled = 2,
                                              axis_pos_x = -1.1, 
                                              axis_pos_y = -1.1, 
                                              threshold = 0.05))

five_gau_proj_tsne_model2 <- plot_proj(
  proj_obj = proj_obj2,
  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
  plot_limits = c(-1.3, 1.3), 
  title = "a3", 
  cex = 1.5,
  position = c(0.08, 0.9),
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio = 1,
        legend.position = "none")
```

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)

## Second projection
projection <- cbind(
  c(0.30943,-0.08452,-0.23005,-0.05284),
  c(0.22463,-0.06682,0.32106,0.02452))

proj_obj2 <- get_projection(projection = projection, 
                            proj_scale = 3, 
                            scaled_data = scaled_gau_data, 
                            scaled_data_model = scaled_gau_data_model, 
                            distance_df_small_edges = distance_df_small_edges, 
                            axis_param = list(limits = 1.2,
                                              axis_scaled = 2,
                                              axis_pos_x = -1.1, 
                                              axis_pos_y = -1.1, 
                                              threshold = 0.05))

five_gau_proj_umap_model2 <- plot_proj(
  proj_obj = proj_obj2,
  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
  plot_limits = c(-1.3, 1.3), 
  title = "b3", 
  cex = 1.5,
  position = c(0.08, 0.9),
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio = 1,
        legend.position = "none")
```

@@

## Second projection
projection <- cbind(
  c(0.30943,-0.08452,-0.23005,-0.05284),
  c(0.22463,-0.06682,0.32106,0.02452))

proj_obj2 <- get_projection(projection = projection, 
                            proj_scale = 3, 
                            scaled_data = scaled_gau_data, 
                            scaled_data_model = scaled_gau_data_model, 
                            distance_df_small_edges = distance_df_small_edges, 
                            axis_param = list(limits = 1.2,
                                              axis_scaled = 2,
                                              axis_pos_x = -1.1, 
                                              axis_pos_y = -1.1, 
                                              threshold = 0.05))

five_gau_proj_pacmap_model2 <- plot_proj(
  proj_obj = proj_obj2,
  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
  plot_limits = c(-1.3, 1.3), 
  title = "c3", 
  cex = 1.5,
  position = c(0.08, 0.9),
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio = 1,
        legend.position = "none")

@@

trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
  five_gau_proj_tsne_model2 +
  five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
  plot_layout(guides = "collect", nrow = 3) &
  theme(legend.position='none')
```
